### 2021-05-09 content/31.tooling-table.tex
Sunday fun with tables in LaTeX

\begin{table}[ht]
	\centering
	\begin{tabular}{llll}
		\toprule
		\textbf{Tool name} & Targets & Language / Platform & Status \\
		\multicolumn{4}{l}{\it Brief Description} \\
		\midrule

		\textbf{Describo} \citep{describo} & Research Data Managers & NodeJS (Desktop) & RC \\
		\multicolumn{4}{l}{\it Interactive desktop application to create, update and export RO-Crates for different profiles} \\

		\textbf{Describo Online} \citep{describo-online} & Platform developers & NodeJS (Web) & Alpha \\
		\multicolumn{4}{l}{\it Web-based application to create RO-Crates using cloud storage} \\

		\textbf{ro-crate-excel} \citep{ro-crate-excel} & Data managers & JavaScript \\
		\multicolumn{4}{l}{\it Command-line tool to help create RO-Crates and HTML-readable rendering} \\

		\textbf{ro-crate-html-js} \citep{ro-crate-html-js} & Developers & JavaScript & Beta \\
		\multicolumn{4}{l}{\it HTML rendering of RO-Crate}  \\

		\textbf{ro-crate-js} \citep{ro-crate-js} & Research Data Managers & JavaScript & Alpha \\
		\multicolumn{4}{l}{\it Library for creating/manipulating crates; basic validation code} \\

		\textbf{ro-crate-ruby} \citep{ro-crate-ruby} & Developers & Ruby & Beta \\
		\multicolumn{4}{l}{\it Ruby library for reading/writing RO-Crate, with workflow support} \\

		\textbf{ro-crate-py} \citep{ro-crate-py} & Developers & Python & Beta \\
		\multicolumn{4}{l}{\it Object-oriented Python library for reading/writing RO-Crate} \\

		\textbf{WorkflowHub} \citep{about-workflowhub} & Workflow users & Ruby & Beta \\
		\multicolumn{4}{l}{\it Workflow repository; imports and exports Workflow RO-Crate} \\

		\textbf{Modern PARADISEC} \citep{modpdsc} & Repository managers & Platform & Beta \\
		\multicolumn{4}{l}{\it Cultural Heritage portal based on OCFL and RO-Crate} \\

		\textbf{ONI express} \citep{arkisto-data-portal} & Repository managers & Platform & Beta \\
		\multicolumn{4}{l}{\it Platform for publishing data and documents stored in an OCFL repository via a web interface} \\

		\textbf{ocfl-tools} \citep{ocfl-tools} & Developers & JavaScript (CLI) & Beta \\
		\multicolumn{4}{l}{\it Tools for managing RO-Crates in an OCFL repository}\\

		\textbf{RO Composer} \citep{ro-composer} & Repository developers & Java & Alpha \\ 
		\multicolumn{4}{l}{\it REST API for gradually building ROs for given profile} \\

		\textbf{galaxy2cwl} \citep{galaxy2cwl} & Workflow developers & Python & Alpha \\
		\multicolumn{4}{l}{\it Wraps Galaxy workflow as Workflow RO-Crate} \\
		\bottomrule
	\end{tabular}
	\caption{Applications and libraries implementing RO-Crate, targeting different types of users across multiple programming languages. Status is indicative as assessed by this work.}
	\label{tab:tools}
\end{table}