### 2021-05-17 latex/ro-crate.tex
Sync

%% ro-crate paper
%% (c) 2020-2021 The University of Manchester
%% 
%% SPDX-License-Identifier: CC-BY-4.0

% add. options: [seceqn,secthm,crcready]
\documentclass[ds]{iosart2x}

%%%%%%%%%%% Put your definitions here

% From https://www.overleaf.com/latex/templates/using-the-stix2-opentype-fonts-with-lualatex-or-xelatex/hdfvhzqmhnpx
\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{hyperref}
\usepackage{fontspec}
\usepackage{unicode-math}
% \usepackage{graphicx}

\DeclareGraphicsExtensions{.pdf,.png,.svg}
\graphicspath{../content/images/}
\usepackage[citations,definitionLists,fencedCode,footnotes,hybrid,pipeTables,smartEllipses,tableCaptions,underscores]{markdown}
\def\markdownOptionOutputDir{../output}

% for table resizing
\usepackage{booktabs}
\usepackage{multirow}

%%%%%%%%%%% End of definitions

\pubyear{2021}
\volume{0}
\firstpage{1}
\lastpage{1}

\begin{document}

% Welcome to the 2000s with UNICODE fonts!
% Adapted from
% https://www.overleaf.com/learn/latex/Articles%2FOpenType-based_math_typesetting:_An_introduction_to_the_STIX2_OpenType_fonts
\setmathfont{STIX2Math}[
    Extension={.otf},
    Path=./STIX2fonts/,
    Scale=1
]
\setmainfont{STIX2Text}[
    Extension={.otf},
    Path=./STIX2fonts/,
    UprightFont={*-Regular},
    BoldFont={*-Bold},
    ItalicFont={*-Italic},
    BoldItalicFont={*-BoldItalic}
]

% https://typeof.net/Iosevka/ mono font has 
% all; the logic symbols we need
\setmonofont{iosevka-slab}[
    Extension={.ttf},
    Path=./ttf-iosevka-slab-6.1.3/,
    Scale=MatchLowercase,
    UprightFont={*-light},
    ItalicFont={*-lightitalic}
]

\begin{frontmatter}

%\pretitle{}
\title{Packaging research artifacts with RO-Crate}
\runtitle{Packaging research artifacts with RO-Crate}
%\subtitle{}

% For one author:
%\author{\inits{N.}\fnms{Name1} \snm{Surname1}\ead[label=e1]{first@somewhere.com}\ead[label=o1,orcid]{0000-0000-0000-0001}}
%\address{Department first, \orgname{University or Company name},
%Abbreviate US states, \cny{Country}\printead[presep={\\}]{e1,o1}}

% Two or more authors:
\author[A,B]{\fnms{Stian} \snm{Soiland-Reyes}\ead[label=e1]{soiland-reyes@manchester.ac.uk}\ead[label=o1,orcid]{0000-0001-9842-9718}%
\thanks{Corresponding author. \printead{e1}.}},
\author[C]{\inits{P.S.}\fnms{Peter} \snm{Sefton}\ead[label=e2]{Peter.Sefton@uts.edu.au}\ead[label=o2,orcid]{0000-0002-3545-944X}},
\author[I]{\inits{M.C.}\fnms{Mercè} \snm{Crosas}\ead[label=e9]{mcrosas@g.harvard.edu}\ead[label=o9,orcid]{0000-0003-1304-1939}},
\author[H]{\inits{D.G.}\fnms{Daniel} \snm{Garijo}\ead[label=e8]{dgarijo@fi.upm.es}\ead[label=o8,orcid]{0000-0003-0454-7145}},
\author[G]{\inits{M.LR.}\fnms{Marco} \snm{La Rosa}\ead[label=e7]{m@lr.id.au}\ead[label=o7,orcid]{0000-}},
\author[E]{\inits{S.L.}\fnms{Simone} \snm{Leo}\ead[label=e5]{simone.leo@crs4.it}\ead[label=o5,orcid]{0000-0001-8271-5429}},
\author[F]{\inits{M.P.}\fnms{Marc} \snm{Portier}\ead[label=e6]{marc.portier@vliz.be}\ead[label=o6,orcid]{0000-0002-9648-6484}},
\author[I]{\inits{A.T.}\fnms{Ana} \snm{Trisovic}\ead[label=e9]{ana.trisovic@gmail.com}\ead[label=o9,orcid]{0000-0003-1991-0533}},

\author[D]{RO-Crate Community},
\author[B]{\inits{P.G.}\fnms{Paul} 
\snm{Groth}\ead[label=e3]{p.t.groth@uva.nl}\ead[label=o3,orcid]{0000-0003-0183-6910}},
\author[A]{\inits{C.G.}\fnms{Carole} \snm{Goble}\ead[label=e3]{carole.goble@manchester.ac.uk}\ead[label=o3,orcid]{0000-0003-1219-2137}}

\address[A]{Computer Science, \orgname{The University of Manchester},
\cny{UK}\printead[presep={\\}]{e1,o1}}
\address[B]{Informatics Institute, \orgname{University of Amsterdam},
\cny{The Netherlands}\printead[presep={\\}]{e2,o2}}
\address[C]{Faculty of Science, \orgname{University Technology Sydney},
\cny{Australia}\printead[presep={\\}]{e3,o3}}
\address[E]{\orgname{Center for Advanced Studies, Research, and Development in Sardinia} (CRS4), Pula (CA), \cny{Italy}\printead[presep={\\}]{e5,o5}}
\address[D]{\url{https://www.researchobject.org/ro-crate/community} (see appendix B)}
\address[H]{Ontology Engineering Group, \orgname{Universidad Politécnica de Madrid},
Madrid, \cny{Spain}\printead[presep={\\}]{e8,o8}}
\address[I]{Institute for Quantitative Social Science, \orgname{Harvard University} Cambridge, MA,  \cny{USA}\printead[presep={\\}]{e9,o9}}

%% TODO: Fix this mess - it seems we need one \address per author
%% as this address also print their individual email and orcid,
%% meaning some organization will repeat.

%\begin{review}{editor}
%\reviewer{\fnms{First} \snm{Editor}\ead[label=oe1,orcid]{}\printead[presep={\ }]{oe1}}
%\reviewer{\fnms{Second} \snm{Editor}\ead[label=oe2,orcid]{}\printead[presep={\ }]{oe2}}
%\end{review}
%\begin{review}{solicited}
%\reviewer{\fnms{First} \snm{Solicited reviewer}\ead[label=os1,orcid]{}\printead[presep={\ }]{os1}}
%\end{review}

\begin{abstract}
\markdownInput{../content/01.abstract.md}
\end{abstract}

\begin{keyword}
\kwd{Data publishing}
\kwd{Data packaging}
\kwd{FAIR}
\kwd{Linked Data}
\kwd{Metadata}
\kwd{Reproducibility}
\kwd{Research Object}
\end{keyword}

\end{frontmatter}

%%%%%%%%%%% The article body starts:

% Every *.md file in content/ in order, except 00, 90
\markdownInput{../content/10.background.md}

\markdownInput{../content/20.rocrate.md}
\markdownInput{../content/21.conceptual.md}
\markdownInput{../content/22.implementation.md}
\markdownInput{../content/23.community.md}

\input{../content/31.tooling-table.tex}
%\markdownInput{../content/31.tooling-table.md}
\markdownInput{../content/30.tooling.md}

\markdownInput{../content/40.inuse.md}
\markdownInput{../content/41.workflows.md}
\markdownInput{../content/42.regulatory.md}
\markdownInput{../content/43.culturalheritage.md}
\markdownInput{../content/44.dmp.md}

\markdownInput{../content/50.relatedwork.md}
\markdownInput{../content/60.conclusion.md}
\markdownInput{../content/70.acknowledgements.md}

%%%%%%%%%%% The bibliography starts:
\nocite{*} 
\bibliographystyle{ios1}
\bibliography{ro-crate}

%%%% Appendix

\appendix
\markdownInput{../content/80.formal-definition.md}
% TODO: Recode formal definition to LaTeX math mode?
%\include{../content/80.formal-definition}

\markdownInput{../content/89.community.md}

\end{document}