### 2021-08-09 output/_markdown_ro-crate/48ea55320d214e1130b767589109006d.md.tex

\markdownRendererHeadingOne{Acknowledgements}\markdownRendererInterblockSeparator
{}\small\markdownRendererInterblockSeparator
{}This work has received funding from the European Commission's Horizon 2020 research and innovation programme for projects \markdownRendererLink{BioExcel-2}{https://cordis.europa.eu/project/id/823830}{https://cordis.europa.eu/project/id/823830}{} (H2020-INFRAEDI-2018-1 823830), \markdownRendererLink{IBISBA}{https://cordis.europa.eu/project/id/730976}{https://cordis.europa.eu/project/id/730976}{} (H2020-INFRAIA-2017-1-two-stage 730976, H2020-INFRADEV-2019-2 871118), \markdownRendererLink{EOSC-Life}{https://cordis.europa.eu/project/id/824087}{https://cordis.europa.eu/project/id/824087}{} (H2020-INFRAEOSC-2018-2 824087), \markdownRendererLink{SyntheSys+}{https://cordis.europa.eu/project/id/823827}{https://cordis.europa.eu/project/id/823827}{} (H2020-INFRAIA-2018-1 823827)\markdownRendererInterblockSeparator
{}A. Trisovic is funded by the Alfred P. Sloan Foundation (grant number P-2020-13988). Harvard Data Commons is supported by an award from Harvard University Information Technology (HUIT).\markdownRendererInterblockSeparator
{}\markdownRendererHeadingTwo{Contributions}\markdownRendererInterblockSeparator
{}Author contributions to this article and the RO-Crate projet according to the Contributor Roles Taxonomy \markdownRendererLink{CASRAI CrEDiT}{https://casrai.org/credit/}{https://casrai.org/credit/}{} \markdownRendererCite{1}+{}{}{doi:10.1087/20150211}:\markdownRendererInterblockSeparator
{}\markdownRendererDlBegin
\markdownRendererDlItem{Stian Soiland-Reyes}\markdownRendererDlDefinitionBegin Conceptualization, Data curation, Formal Analysis, Funding acquisition, Investigation, Methodology, Project administration, Software, Visualization, Writing – original draft, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Peter Sefton}\markdownRendererDlDefinitionBegin Conceptualization, Investigation, Methodology, Project administration, Resources, Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Mercè Crosas}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Leyla Jael Castro}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Frederik Coppens}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{José M. Fernández}\markdownRendererDlDefinitionBegin Methodology, Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Daniel Garijo}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Marco La Rosa}\markdownRendererDlDefinitionBegin Software, Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Simone Leo}\markdownRendererDlDefinitionBegin Software, Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Eoghan Ó Carragáin}\markdownRendererDlDefinitionBegin Investigation, Methodology, Project administration, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Marc Portier}\markdownRendererDlDefinitionBegin Methodology, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Ana Trisovic}\markdownRendererDlDefinitionBegin Software, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{RO-Crate Community}\markdownRendererDlDefinitionBegin Investigation, Software, Validation, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Paul Groth}\markdownRendererDlDefinitionBegin Methodology, Supervision, Writing – original draft, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Carole Goble}\markdownRendererDlDefinitionBegin Conceptualization, Funding acquisition, Methodology, Project administration, Supervision, Visualization, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd 
\markdownRendererDlEnd\markdownRendererInterblockSeparator
{}We would also like to thank for contributions from:\markdownRendererInterblockSeparator
{}\markdownRendererDlBegin
\markdownRendererDlItem{Finn Bacall}\markdownRendererDlDefinitionBegin Software, Methodology\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Herbert Van de Sompel}\markdownRendererDlDefinitionBegin Writing – review \& editing \markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Ignacio Eguinoa}\markdownRendererDlDefinitionBegin Software, Methodology\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Nick Juty}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Oscar Corcho}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Stuart Owen}\markdownRendererDlDefinitionBegin Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd \markdownRendererDlItem{Laura Rodríguez-Navas}\markdownRendererDlDefinitionBegin Software, Visualization, Writing – review \& editing\markdownRendererDlDefinitionEnd \markdownRendererDlItemEnd 
\markdownRendererDlEnd\markdownRendererInterblockSeparator
{}% CrEDiT taxonomy: % % Conceptualization % Data curation % Formal Analysis % Funding acquisition % Investigation % Methodology % Project administration % Resources % Software % Supervision % Validation % Visualization % Writing – original draft % Writing – review \& editing\relax